---
title: "R Notebook"
output: html_notebook
---

# Consider sales data, perform Exploratory Data Analysis (EDA) and Feature Engineering

using appropriate techniques and variables

## The data set

```{r}
library(readxl)
Sales_Data <- read_excel("Sales_Data.xlsx")
View(Sales_Data)
head(Sales_Data)
```

## Summary of the data set

```{r}
summary(Sales_Data)
```

## Missing values in each column

```{r}
missing_data = data.frame(column = colnames(Sales_Data),
                     missing_val = colSums(Sales_Data == "NULL!"),
                     percentage_missing = colMeans(Sales_Data == "NULL!") * 100)
View(missing_data)
missing_data

```

```{r}
# check for na 
any(is.na(Sales_Data))
```

```{r}
# check for null
any(is.null(Sales_Data))
```

The data has missing values represented as `Null!` and not any of the default `NA` and `NULL`

## Check the type of data 

```{r}
str(Sales_Data)
```

## Convert the data to the numeric values

```{r}
Sales_Data%in% seq3:15,] = as.numeric(Sales_Data[3:15,])

str(Sales_Data)
```

## Number of data points

```{r}
ncol(Sales_Data)
```

```{r}
nrow(Sales_Data)
```

```{r}
ncol(Sales_Data) * ncol(Sales_Data)
```

There are `225` data points represented as `15` columns and `157` rows

## Unique manufacturers

```{r}
unique(Sales_Data$manufact)
```

There are `31` unique car manufacturers

## Cars produced by each manufacturer

```{r}
cars = data.frame(manufact = Sales_Data$manufact,
                  model = Sales_Data$model,
                  count = rep(1, 157))
library(dplyr)
cars = cars %>% group_by(manufact) %>%summarise(count = sum(count))
```

```{r}
library(ggplot2)

ggplot(cars, aes(x= "", y = count, fill = manufact))+
  geom_bar(stat = "identity", width = 1)+
  coord_polar("y", start = 0)+
  geom_text(aes(label = paste0(count, "%")), position = position_stack(vjust=0.5))+
  theme_classic() +
  theme(axis.line = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank())
```
