---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

# Consider sales data, perform Exploratory Data Analysis (EDA) and Feature Engineering

using appropriate techniques and variables

## The data set

```{r}
library(readxl)
Sales_Data <- read_excel("Sales_Data.xlsx")
View(Sales_Data)
head(Sales_Data)
```

## Summary of the data set

### Columns of the data set

```{r}
colnames(Sales_Data)
```

### Convert the number column data to the numeric values

```{r}
numeric_cols = c("sales","resale","type","price", "engine_s", "horsepow",  "wheelbas",  "width", "length", "curb_wgt", "fuel_cap","mpg",  "Grp_Horse")
Sales_Data[, numeric_cols] = lapply(Sales_Data[,numeric_cols], as.numeric)  
str(Sales_Data)
```

### Data set distribution

```{r}
summary(Sales_Data)
```

## Missing values in each column

```{r}
num_col = Sales_Data[,numeric_cols]
str_col = Sales_Data[,1:2]

missing_data = data.frame(columns = colnames(str_col),
                          missing_val = colSums(str_col == "NULL"),
                          perc_missing = colMeans(str_col == "NULL"))
num_miss = data.frame(columns = colnames(num_col),
                          missing_val = colSums(is.na(num_col)),
                          perc_missing = colMeans(is.na(num_col)))

missing_data = rbind(missing_data, num_miss)
View(missing_data)
missing_data

```

```{r}
# check for na 
any(is.na(Sales_Data))
```

```{r}
# check for null
any(is.null(Sales_Data))
```

The data has missing values represented as `Null!` and not any of the default `NA` and `NULL`

The data has no column with more than half the values in the column missing so no columns or rows need to be dropped

## Number of data points

```{r}
ncol(Sales_Data)
```

```{r}
nrow(Sales_Data)
```

```{r}
ncol(Sales_Data) * ncol(Sales_Data)
```

There are `225` data points represented as `15` columns and `157` rows

## Unique manufacturers

### Impute the NULL manufacturers

This is done by imputing the previously observed value in the column

```{r}
library(zoo)
Sales_Data$manufact[Sales_Data$manufact == "NULL!"] = NA
Sales_Data$manufact = na.locf(Sales_Data$manufact)

# confirm if any na are in the manufact column 
any(is.na(Sales_Data$manufact))
```

### The number of manufacturers

```{r}
unique(Sales_Data$manufact)
```

There are `30` unique car manufacturers

## Cars produced by each manufacturer

```{r}
cars = data.frame(manufact = Sales_Data$manufact,
                  model = Sales_Data$model,
                  count = rep(1, 157))
library(dplyr)
cars = cars %>% group_by(manufact) %>%summarise(count = sum(count))
```

### See the cars produced

```{r}
cars
```

```{r}
library(ggplot2)

ggplot(cars, aes(x= "", y = count, fill = manufact))+
  geom_bar(stat = "identity", width = 1)+
  coord_polar("y", start = 0)+
  theme_classic() +
  theme(axis.line = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank())
```

## Assessing the numeric columns

```{r}
colnames(Sales_Data[,numeric_cols])
```

### Separate factor and number columns

**Discrete/ Factor Columns**

1.  Type
2.  Grp_Horse
3.  engine_s
4.  horsepow
5.  mpg

These are imputed with the modal value

### Continuous columns

1.  sales
2.  resale
3.  price
4.  wheelbas
5.  width
6.  length
7.  curb_wgt
8.  fuel_cap

### imputation of discrete / factor using mode

```{r}
# distribution plot function
plot_distrib = function(col_name){
  # convert the data into columns
  col_vals = data.frame(table(col_name))
  col_vals$percent = col_vals$Freq / sum(col_vals$Freq) * 100
  # plot the distibution
  plot = ggplot(col_vals, aes(x = "", y = Freq, fill = as.factor(as.character(col_name)))) +
    geom_bar(stat = "identity") +
    coord_polar("y") +
    geom_text(aes(label = paste0(round(percent), "%")), position = position_stack(vjust = 0.5)) +
    ggtitle("Distribution Percentages") +
    theme_light()
  return(plot)
}
```

```{r}
# imputation func
calc_mod = function(x){
  uniq_x = unique(x)
  uniq_x[which.max(tabulate(match(x, uniq_x)))]
}
```

1.  **Type**

```{r}
plot_distrib(Sales_Data$type)
Sales_Data$type = ifelse(is.na(Sales_Data$type), calc_mod(Sales_Data$type), Sales_Data$type)

# Check if any na is left in the column
if(any(is.na(Sales_Data$type))){
  print("There are still na values in: type")
}else{
  print("No na values in: type")
}

# replot the distribution to see if it is consistent
plot_distrib(Sales_Data$type)
```

2.  **Grp_Hourse**

```{r}
plot_distrib(Sales_Data$Grp_Horse)
Sales_Data$Grp_Horse = ifelse(is.na(Sales_Data$Grp_Horse), calc_mod(Sales_Data$Grp_Horse), Sales_Data$Grp_Horse)

# Check if any na is left in the column
if(any(is.na(Sales_Data$Grp_Horse))){
  print("There are still na values in: Grp_Horse")
}else{
  print("No na values in: Grp_Horse")
}

# replot the distribution to see if it is consistent
plot_distrib(Sales_Data$Grp_Horse)
```

3.  **engine_s**

```{r}
plot_density = function(data_set,num_col){
  ggplot(data_set, aes(x = data_set[[num_col]]))+
    geom_density(fill = "lightblue", color = "blue")+
    ggtitle(paste("Density Plot of", num_col))+
    xlab(num_col)+
    ylab("Density")+
    theme_minimal()
}
```

```{r}
plot_density(Sales_Data,"engine_s")
Sales_Data$engine_s = ifelse(is.na(Sales_Data$engine_s), calc_mod(Sales_Data$engine_s), Sales_Data$engine_s)

# Check if any na is left in the column
if(any(is.na(Sales_Data$engine_s))){
  print("There are still na values in: engine_s")
}else{
  print("No na values in: engine_s")
}

# replot the distribution to see if it is consistent
plot_density(Sales_Data,"engine_s")
```

4.  **horsepow**

```{r}

plot_density(Sales_Data,"horsepow")
Sales_Data$horsepow = ifelse(is.na(Sales_Data$horsepow), calc_mod(Sales_Data$horsepow), Sales_Data$horsepow)

# Check if any na is left in the column
if(any(is.na(Sales_Data$horsepow))){
  print("There are still na values in: horsepow")
}else{
  print("No na values in: horsepow")
}

# replot the distribution to see if it is consistent
plot_density(Sales_Data,"horsepow")
```

```{r}
plot_density(Sales_Data,"mpg")
Sales_Data$mpg = ifelse(is.na(Sales_Data$mpg), calc_mod(Sales_Data$mpg), Sales_Data$mpg)

# Check if any na is left in the column
if(any(is.na(Sales_Data$mpg))){
  print("There are still na values in: mpg")
}else{
  print("No na values in: mpg")
}

# replot the distribution to see if it is consistent
plot_density(Sales_Data,"mpg")
```

### Imputing continuous columns

using a random value based on the number of na values in the column

```{r}
set.seed(123)
rand_val = function(data, column_name, low, up){
  data[[column_name]] = ifelse(is.na(data[[column_name]]), runif(sum(is.na(data[[column_name]])), min = low, max = up), data[[column_name]])
  return(data)
}
```

1 **sales**

```{r}
plot_density(Sales_Data, "sales")
Sales_Data=rand_val(Sales_Data, "sales", 
             min(Sales_Data$sales, na.rm = TRUE),
             max(Sales_Data$sales, na.rm = TRUE))

if(any(is.na(Sales_Data$sales))){
  print(paste("there are na values", "sales"))
}else{
  print(paste("no na values", "sales"))
}
plot_density(Sales_Data, "sales")
```

2.  **resale**

```{r}
plot_density(Sales_Data, "resale")
Sales_Data=rand_val(Sales_Data, "resale", 
             min(Sales_Data$resale, na.rm = TRUE),
             max(Sales_Data$resale, na.rm = TRUE))

if(any(is.na(Sales_Data$resale))){
  print(paste("there are na values", "resale"))
}else{
  print(paste("no na values", "resale"))
}
plot_density(Sales_Data, "resale")
```

3.  **price**

```{r}
low = min(Sales_Data$price, na.rm = TRUE)
up = max(Sales_Data$price, na.rm = TRUE)
plot_density(Sales_Data, "price")
Sales_Data=rand_val(Sales_Data, "price",low,up)

if(any(is.na(Sales_Data$price))){
  print(paste("there are na values", "price"))
}else{
  print(paste("no na values", "price"))
}
plot_density(Sales_Data, "price")
```

4.  **wheelbas**

```{r}
low = min(Sales_Data$wheelbas, na.rm = TRUE)
up = max(Sales_Data$wheelbas, na.rm = TRUE)
plot_density(Sales_Data, "wheelbas")
Sales_Data=rand_val(Sales_Data, "wheelbas",low,up)

if(any(is.na(Sales_Data$wheelbas))){
  print(paste("there are na values", "wheelbas"))
}else{
  print(paste("no na values", "wheelbas"))
}
plot_density(Sales_Data, "wheelbas")
```

5.  **width**

```{r}
low = min(Sales_Data$width, na.rm = TRUE)
up = max(Sales_Data$width, na.rm = TRUE)
plot_density(Sales_Data, "width")
Sales_Data=rand_val(Sales_Data, "width",low,up)

if(any(is.na(Sales_Data$width))){
  print(paste("there are na values", "width"))
}else{
  print(paste("no na values", "width"))
}
plot_density(Sales_Data, "width")
```

6.  **length**

```{r}
low = min(Sales_Data$length, na.rm = TRUE)
up = max(Sales_Data$length, na.rm = TRUE)
plot_density(Sales_Data, "length")
Sales_Data=rand_val(Sales_Data, "length",low,up)

if(any(is.na(Sales_Data$length))){
  print(paste("there are na values", "length"))
}else{
  print(paste("no na values", "length"))
}
plot_density(Sales_Data, "length")
```

7.  **curb_wgt**

```{r}
low = min(Sales_Data$curb_wgt, na.rm = TRUE)
up = max(Sales_Data$curb_wgt, na.rm = TRUE)
plot_density(Sales_Data, "curb_wgt")
Sales_Data=rand_val(Sales_Data, "curb_wgt",low,up)

if(any(is.na(Sales_Data$curb_wgt))){
  print(paste("there are na values", "curb_wgt"))
}else{
  print(paste("no na values", "curb_wgt"))
}
plot_density(Sales_Data, "curb_wgt")
```

8.  **fuel_cap**

```{r}
low = min(Sales_Data$fuel_cap, na.rm = TRUE)
up = max(Sales_Data$fuel_cap, na.rm = TRUE)
plot_density(Sales_Data, "fuel_cap")
Sales_Data=rand_val(Sales_Data, "fuel_cap",low,up)

if(any(is.na(Sales_Data$fuel_cap))){
  print(paste("there are na values", "fuel_cap"))
}else{
  print(paste("no na values", "fuel_cap"))
}
plot_density(Sales_Data, "fuel_cap")
```

# Data Visuals

## spread of price

```{r}
ggplot(Sales_Data, aes(x = Sales_Data$price)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
  labs(x = "Price", y = "Frequency", title = "Histogram of Price")
```

-   price has a `right skew`

-   This means that most cars in the sales data have a `relatively low prices`

## Engine size and sales

```{r}
engine_data = Sales_Data[, c(7, 3, 13, 8, 14)]
engine_data$count = rep(1, nrow(Sales_Data))
engine_data = engine_data %>% group_by(engine_s)%>%summarise(
  sales = sum(sales),
  av_fuel_cap = mean(fuel_cap),
  av_horsepow = mean(horsepow),
  av_mpg = mean(mpg)
)
engine_data

```

This data represents the `total sales`, `average horsepower`, `average mpg` and `average fuel capacity` for each engine size

## Regression Analysis

### Variables that influence horse power

```{r}
# linear regression data
cols_lin_regress = c(3, 4, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15)

# 
linear_regress_data = Sales_Data[, cols_lin_regress]
colnames(linear_regress_data)
```

```{r}
linear_model = lm(horsepow~., data = linear_regress_data)
summary(linear_model)
```

```{r}
linear_model_2 = lm(horsepow~Grp_Horse+engine_s+price, data = linear_regress_data)
summary(linear_model_2)
```

From the data the `horse power` can be predicted based on `Grp horse`, `engine size`, and `price`

## Homoscedasticity

```{r}
library(lmtest)
homoscedasticity_test = bptest(linear_model_2)
homoscedasticity_test
```

The p-value `0.03414` is less than `0.05` indicates that the data set has a constant variance hence the data fits `homoscedascity`

## Normality Test

```{r}
linear_residuals = residuals(linear_model_2)

# qqplot
qqnorm(linear_residuals)
qqline(linear_residuals, col = 2)
```

Most of the points lie on the qqline indicating that the variables follow a normal distribution

## linearity

```{r}
library(ggplot2)

res_fit = data.frame(fitted = linear_model_2$fitted.values,
                     residual = linear_model_2$residuals)

ggplot(data = res_fit,
       aes(x = fitted, y = residual)) +
  geom_point()+
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(
    x = "fitted values",
    y = "residuals",
    title = "Fitted value Residual Plot"
  )
```

Most data points are around the `0` line and equally distributed either side meaning that the relationship of the dependent and independent variables is `linear`

## multi-colinearity

```{r}
library(car)
var_inf= vif(linear_model_2)
var_inf
```

The variance inflation factor for each of the variables is `less than 5` indicating that the variables are not affected by variance inflation due to **multi-colinearity**

# Factor analysis

## Corelation

```{r}
pca_data = linear_regress_data
pca_data = pca_data %>% select_if(is.numeric)
data_pca_cor = cor(pca_data)
data_pca_cor
```

## Principle components

```{r}
pca_result = prcomp(pca_data, scale = TRUE) # std dev = 1 and mean = 0
summary(pca_result)
```

-   **Standard deviation:** Indicates the amount of variance captured by each principal component. Larger standard deviations mean that the principal component captures more variability in the data.

-   **Proportion of Variance:** Represents the proportion of total variance in the data explained by each principal component. Higher values indicate that the corresponding principal component explains more of the total variance.

-   **Cumulative Proportion:** Shows the cumulative proportion of variance explained by each principal component. It demonstrates how much of the total variance is explained when considering up to that specific component.

Therefore

`pc1` explains `49.23%` of the total variation

`pc2` explains `67.92%`

...

## Scree plot

```{r}
library(psych)
library(FactoMineR)
library(factoextra)
fviz_eig(pca_result, 
         addlabels = TRUE, 
         ylim = c(0, 70))
```

Taking 3 principle components will ensure the highest explained variance

## Key components

```{r}
pca_result$rotation
```

## Transform based on the principle components

```{r}
transformed = predict(pca_result, newdata =linear_regress_data)
head(transformed)
```

It is clear that `pc12` experiences the least variance captured by the dimension reduction

## Data point labels

```{r}
linear_regress_data$mpg = ifelse(linear_regress_data$mpg < 21, 'Guzzler', "Fuel Efficient")
linear_regress_data$mpg = as.factor(linear_regress_data$mpg)
fviz_pca_biplot(pca_result,
                label = "var",
                habillage = linear_regress_data$mpg)
```

The higher the `mpg` the more likely the car is fuel efficient

```{r}
colnames(linear_regress_data)
```

`` resale, price, engine size, horsepow` `` increase with a decrease in fuel efficiency

`wheelbas, width, length, curb weight, fuel capacity and grp horse` decrease with an increase in the car being a guzzler

`sales` has a negative non impact to being a guzzler or not
